---
title: "Classnotes"
author: "Gibi Thomas"
date: "2026-04-02"
output:
  word_document: default
  html_document: default
---

1. 1. Give me notes on the recap material. These questions focus on material that students commonly struggle with (data wrangling, plotting, modeling, or debugging). 
```{r}
```{r}
library(tidyverse)
```

```{r}
microbiome.fungi <- read.csv("C:/Users/My PC/OneDrive/문서/PLPA6820/Bull_richness.csv")
str(microbiome.fungi)
```


#### `select()`
```{r}
microbiome.fungi2 <- select(microbiome.fungi, SampleID, Crop, Compartment:Fungicide, richness)
str(microbiome.fungi2)
```


#### `filter()`

```{r}
head(filter(microbiome.fungi2, Treatment == "Conv."))

head(filter(microbiome.fungi2, Treatment == "Conv." & Fungicide == "C"))

head(filter(microbiome.fungi2, Sample == "A" | Sample == "B")) # samples A or B
```

#### `mutate()`

```{r}
microbiome.fungi2$logRich <- log(microbiome.fungi2$richness) 
# Create a new column called logRich
head(mutate(microbiome.fungi2, logRich = log(richness)))

# Creating a new column which combines Crop and Treatment 
head(mutate(microbiome.fungi2, Crop_Treatment = paste(Crop, Treatment)))
```

#### the pipe `%>%`

```{r}

select(microbiome.fungi, SampleID, Crop, Compartment:Fungicide, richness) # before

microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns
  filter(Treatment == "Conv.") %>% # subsetting to only include the conventional treatment
  mutate(logRich = log(richness)) %>% # creating a new column of the log richness
  head() # displaying the first six rows
```

#### `summarise()`

```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns
  filter(Treatment == "Conv.") %>% # subsetting to only include the conventional treatment
  mutate(logRich = log(richness)) %>% # creating a new column of the log richness
  summarise(Mean.rich = mean(logRich)) # calculating overall mean log richness within the conventionally managed treatment
```

 
```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns
  filter(Treatment == "Conv.") %>% # subsetting to only include the conventional treatment
  mutate(logRich = log(richness)) %>% # creating a new column of the log richness
  summarise(Mean.rich = mean(logRich), # calculating the mean richness, stdeviation, and standard error
            n = n(), 
            sd.dev = sd(logRich)) %>%
  mutate(std.err = sd.dev/sqrt(n))
```

 
```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  mutate(logRich = log(richness)) %>% # creating a new column of the log richness
  summarise(Mean.rich = mean(logRich), # calculating the mean richness, stdeviation, and standard error
            n = n(), 
            sd.dev = sd(logRich)) %>%
  mutate(std.err = sd.dev/sqrt(n))
```

#### Connecting to plotting

```{r}
plot1 <- microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  mutate(logRich = log(richness)) %>% # creating a new column of the log richness
  summarise(Mean.rich = mean(logRich), # calculating the mean richness, stdeviation, and standard error
            n = n(), 
            sd.dev = sd(logRich)) %>%
  mutate(std.err = sd.dev/sqrt(n)) %>%
  ggplot(aes(x = Fungicide, y = Mean.rich)) + # adding in a ggplot
  geom_bar(stat="identity") +
  geom_errorbar( aes(x=Fungicide, ymin=Mean.rich-std.err, ymax=Mean.rich+std.err), width=0.4) +
  theme_minimal() +
  xlab("") +
  ylab("Log Richness") +
  facet_wrap(~Treatment)
```

#### Joining 

```{r}
# selecting just the richness and sample ID
richness <- microbiome.fungi %>%
  select(SampleID, richness)

# selecting columns that don't include the richness
metadata <- microbiome.fungi %>% 
  select(SampleID, Fungicide, Crop, Compartment, GrowthStage, Treatment, Rep, Sample)

head(metadata)
head(richness)


head(left_join(metadata, richness, by = "SampleID")) # adding the richness data to the metadata based on on the common column of sampleID
```


#### Pivoting

```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% 
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  summarise(Mean = mean(richness)) # calculates the mean per Treatment and Fungicide 
```


```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns  
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  summarise(Mean = mean(richness)) %>% # calculates the mean per Treatment and Fungicide 
  pivot_wider(names_from = Fungicide, values_from = Mean) # pivot to wide format
```

 
```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns  filter(Class == "Sordariomycetes") %>%
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  summarise(Mean = mean(richness)) %>% # calculates the mean per Treatment and Fungicide 
  pivot_wider(names_from = Fungicide, values_from = Mean) %>% # pivot to wide format
  mutate(diff.fungicide = C - F) # calculate the difference between the means. 
```


```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns  filter(Class == "Sordariomycetes") %>%
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  summarise(Mean = mean(richness)) %>% # calculates the mean per Treatment and Fungicide 
  pivot_wider(names_from = Fungicide, values_from = Mean) %>% # pivot to wide format
  mutate(diff.fungicide = C - F) %>%  # calculate the difference between the means. 
  ggplot(aes(x = Treatment, y = diff.fungicide)) + # Plot it 
  geom_col() +
  theme_minimal() +
  xlab("") +
  ylab("Difference in average species richness")
```
```
2. 2. Complete one of the extra optional materials provided, OR explore a topic of your choice, or find a topic online and explore the code and how it works.

Write a short justification (2–3 sentences):
Why did you choose this? What did you want to learn?

I wanted to try a forest plot as it helps visualize both the magnitude and uncertainty of responses in a compact and interpretable format. 
```{r}
library(dplyr)
library(ggplot2)
library(emmeans)
```


```{r}
setwd("C:/Users/My PC/OneDrive/문서/PLPA6820")

# Read CSV
df <- read.csv("RF1_8.csv")

# Model
model <- lm(RV ~ CC * Biochar * Cycle, data = df)  # <- use 'df'

# Estimated means
library(emmeans)
emm <- emmeans(model, ~ CC * Biochar * Cycle)
emm_df <- as.data.frame(emm)

# Clean data
library(dplyr)
emm_df <- emm_df %>%
  mutate(
    Cycle = factor(Cycle, levels = paste0("R", 1:8)),
    CC_Biochar = paste(CC, Biochar, sep = "_")
  )

# Forest-style plot
library(ggplot2)
ggplot(emm_df, aes(x = emmean, y = Cycle)) +
  geom_point(aes(color = Biochar, shape = CC),
             position = position_dodge(width = 0.6),
             size = 3) +
  geom_errorbarh(aes(xmin = lower.CL, xmax = upper.CL, color = Biochar),
                 position = position_dodge(width = 0.6),
                 height = 0.2) +
  theme_bw() +
  labs(
    title = "Forest Plot: CC × Biochar across Rainfall Cycles",
    x = "RV",
    y = "Rainfall Cycle"
  ) +
  facet_wrap(~ CC) +
  geom_vline(xintercept = mean(emm_df$emmean), linetype = "dashed")

```




3. Introduction to renv

renv: instead of storing the package into your system it stores everything in your R project. locked file will be put up in your GitHub. 
```{r}
library(ggplot2)
library(dplyr)
```
```{r}
install.packages("renv")
```
```{r}

```
library directory appears on the files which tells us that whatever we install everything is in there.

source("renv/activate.R")

```{r}
install.packages("MASS")
```


when someone else tries to install the packages that was pushed to the GitHub its shows error because rnev updates the functions to install a particular package not the package itself. 

